Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-18 doc/GMP2015/s01_introduction.tex

\newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}

Some properties of natural and engineered materials can be analyzed by
treating them as stacked two dimensional (2D) layers. The structure
within each layer could be self-similar \cite{Intro1} spanning
multiple scales, generally aperiodic and quasi-uniform within any one
scale, and consisting of a few repeated motifs appearing in disordered
arrangements. Each layer is not necessarily planar, i.e., it consists
of multiple, inter-constraining planar (genus 0) monolayers.
Furthermore, a layer is often  either {\sl isostatic or
under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
definitions). These properties are consistent with a self-assembled 2D
structure that minimizes mass, optimally distributes external stresses
and itself participates in the assembly of diverse and
multifunctional, larger 2D structures.

@@

    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \cite{XX}

@@

    \item Silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
\end{enumerate}
%
In order to study structural and mechanical properties such as
rigidity, flexes and distribution of external stresses within the
material layer, it is natural to model a material layer as a solution
of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D {\bf{\em
qusecs}} (quasi-uniform or self-similar constraint systems) can be
used to understand or design material layers (their solutions) with
desired properties.

\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
\textit{qusecs},  many of which are \textit{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
\textit{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \textit{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

\meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
locating all maximal, generically rigid subsystems \cite{XX}. When the
input is rigid, these algorithms compute the identity and output the
input.

@@

\meeraheader{ii}{Optimal Recursive Decomposition (DR-planning)}
Recursive decomposition of geometric constraint systems has been
formalized \cite{XX} and well-studied \cite{XX} as the {\sl
Decomposition-Recombination (DR)-planning} problem. For the above-
mentioned classes of 2D qusecs, generic rigidity is a combinatorial
property and hence each level of the decomposition should, in
principle, be achievable by a graph-based algorithm as in (1), without
involving geometric information in the constraint system. Since many
such decompositions can exist for a given constraint system, criteria
defining desirable or optimal DR-plans and DR-planning algorithms were
given in \cite{XX}. An \emph{optimal DR-plan} is one that minimizes
the maximum number of child subsystems of any parent system, a
property that is crucial for obtaining a solution to the parent
constraint system from the solutions of the child systems.

However, for general 2D qusecs, even when restricted to bar-joint
systems, the optimal DR-planning problem was shown to be NP-hard
\cite{XX}.

\meeraheader{iii}{DR-plans for special classes and with other
criteria}
For a special class of 2D qusecs, namely \emph{tree- decomposable}
systems \cite{XX} common in computer aided mechanical design, (which
includes ruler-and-compass and Henneberg-I constructible systems), all
DR-plans turn out to be optimal. This satisfies the so-called \emph
{Church-Rosser} criterion, leading to highly efficient DR-planning
algorithms. For general 2D qusecs, alternate criteria were suggested
such as \emph{cluster minimality} requiring parent systems to be
composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
no proper subset forms a rigid system); and \emph{proper maximality},
requiring child subsystems to be maximal rigid proper subsystems of
the parent system.

While polynomial time algorithms were given \cite{XX} to generate DR-
plans meeting the cluster minimality criterion, no such algorithm is
known for the latter criterion.

\meeraheader{iv}{Optimal Recombination and Solution Space Navigation}
For the entire DR-plan, finding all desired solutions is barely
tractable even if recombination of solved subsystems is easy for each
indecomposable parent system in the DR-plan. This is because even for
the simplest, highly decomposable systems with small DR-plans, the
problem of finding even a single solution to the input system at the
root of the DR-plan is NP-hard \cite{XX} and there is a combinatorial
explosion of solutions \cite{XX}. Typically, however, the desired
solution has a given orientation type, in which case, the crux of the
difficulty is concentrated in the algebraic complexity of
(re)combining child system solutions, to give a solution to the parent
system at any given level of the DR-plan. For fairly general 3D
constraint systems, there are algorithms with formal guarantees that
uncover underlying matroids to combinatorially obtain an optimal
parameterization to minimize the algebraic complexity of the
indecomposable parent (sub)systems that occur in the DR-plan \cite{XX,
XX,XX}, provided the DR-plan meets some of the abovementioned
criteria.

However, the generality of these algorithms trades-off against
efficiency, and despite the optimization, the best algorithms can
still take exponential time in the number of child subsystems (which
can be arbitrarily large even for optimal DR-plans) in order to
guarantee all solutions of a given orientation type, even for a single
(sub)system in a DR-plan. They are prohibitively slow in practice. We
note that, utilizing the DR-plan and optimal recombination as a
principled basis, high performance heuristics and software exists
\cite{XX,XX} to tame combinatorial explosion via user intervention.

\meeraheader{v}{Configuration Spaces of Under-constrained Systems}
For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \cite{XX, XX} to complete them into isostatic systems
\cite{XX, XX} and to find paths within the connected components
\cite{XX,XX} of standard Cartesian configuration spaces. Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \cite{XX,XX} to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist \cite{XX,XX}, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.

@@

The contributions of this paper are the following.
\begin{itemize}
  \item We navigate the NP-hardness barrier mentioned in (2) above,
  for finding optimal DR-plans by defining a so-called \textit{canonical}
  DR-plan and showing a strong Church-Rosser property: \textit{all
  canonical DR-plans for isostatic or under-constrained 2D qusecs are
  optimal} (Section \ref{sec:DRP}).

@@

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system. Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems \cite{XX}. When the required
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in \cite{XX,XX} (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems \cite{XX})
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems).

>>>>

% \newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}

Some properties of natural and engineered materials can be analyzed by
treating them as stacked two dimensional (2D) layers. The structure
within each layer could be self-similar \cite{Intro1} spanning
multiple scales, generally aperiodic and quasi-uniform within any one
scale, and consisting of a few repeated motifs appearing in disordered
arrangements. Each layer is not necessarily planar, i.e., it consists
of multiple, inter-constraining planar (genus 0) monolayers.
Furthermore, a layer is often  either \vemph{isostatic or
under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
definitions). These properties are consistent with a self-assembled 2D
structure that minimizes mass, optimally distributes external stresses
and itself participates in the assembly of diverse and
multifunctional, larger 2D structures.

@@

    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited

@@

    \item Silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
\end{enumerate}
%
In order to study structural and mechanical properties such as
rigidity, flexes and distribution of external stresses within the
material layer, it is natural to model a material layer as a solution
of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D
\dfn{qusecs} (quasi-uniform or self-similar constraint systems) can be
used to understand or design material layers (their solutions) with
desired properties.

\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
\dfn{qusecs},  many of which are \dfn{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
\dfn{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

\medskip\noindent
\header{(i) Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
locating all maximal, generically rigid subsystems \uncited. When the
input is rigid, these algorithms compute the identity and output the
input.

@@

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning)}
Recursive decomposition of geometric constraint systems has been
formalized \uncited and well-studied \uncited as the \dfn
{Decomposition-Recombination (DR)-planning} problem. For the above-
mentioned classes of 2D qusecs, generic rigidity is a combinatorial
property and hence each level of the decomposition should, in
principle, be achievable by a graph-based algorithm as in (1), without
involving geometric information in the constraint system. Since many
such decompositions can exist for a given constraint system, criteria
defining desirable or optimal DR-plans and DR-planning algorithms were
given in \uncited. An \dfn{optimal DR-plan} is one that minimizes
the maximum number of child subsystems of any parent system, a
property that is crucial for obtaining a solution to the parent
constraint system from the solutions of the child systems.

However, for general 2D qusecs, even when restricted to bar-joint
systems, the optimal DR-planning problem was shown to be NP-hard
\uncited.

\medskip\noindent
\header{(iii) DR-plans for special classes and with other
criteria}
For a special class of 2D qusecs, namely \dfn{tree-decomposable}
systems \uncited common in computer aided mechanical design, (which
includes ruler-and-compass and Henneberg-I constructible systems), all
DR-plans turn out to be optimal. This satisfies the so-called \dfn
{Church-Rosser} criterion, leading to highly efficient DR-planning
algorithms. For general 2D qusecs, alternate criteria were suggested
such as \dfn{cluster minimality} requiring parent systems to be
composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
no proper subset forms a rigid system); and \dfn{proper maximality},
requiring child subsystems to be maximal rigid proper subsystems of
the parent system.

While polynomial time algorithms were given \uncited to generate DR-
plans meeting the cluster minimality criterion, no such algorithm is
known for the latter criterion.

\medskip\noindent
\header{(iv) Optimal Recombination and Solution Space Navigation}
For the entire DR-plan, finding all desired solutions is barely
tractable even if recombination of solved subsystems is easy for each
indecomposable parent system in the DR-plan. This is because even for
the simplest, highly decomposable systems with small DR-plans, the
problem of finding even a single solution to the input system at the
root of the DR-plan is NP-hard \uncited and there is a combinatorial
explosion of solutions \uncited. Typically, however, the desired
solution has a given orientation type, in which case, the crux of the
difficulty is concentrated in the algebraic complexity of
(re)combining child system solutions, to give a solution to the parent
system at any given level of the DR-plan. For fairly general 3D
constraint systems, there are algorithms with formal guarantees that
uncover underlying matroids to combinatorially obtain an optimal
parameterization to minimize the algebraic complexity of the
indecomposable parent (sub)systems that occur in the DR-plan \uncited\uncited\uncited, provided the DR-plan meets some of the abovementioned
criteria.

However, the generality of these algorithms trades-off against
efficiency, and despite the optimization, the best algorithms can
still take exponential time in the number of child subsystems (which
can be arbitrarily large even for optimal DR-plans) in order to
guarantee all solutions of a given orientation type, even for a single
(sub)system in a DR-plan. They are prohibitively slow in practice. We
note that, utilizing the DR-plan and optimal recombination as a
principled basis, high performance heuristics and software exists
\uncited\uncited to tame combinatorial explosion via user intervention.

\medskip\noindent
\header{(v) Configuration Spaces of Under-constrained Systems}
For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \uncited\uncited to complete them into isostatic systems
\uncited\uncited and to find paths within the connected components
\uncited\uncited of standard Cartesian configuration spaces. Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \uncited\uncited to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist \uncited\uncited, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.

@@

The contributions of this paper are the following.
\begin{itemize}
  \item We navigate the NP-hardness barrier mentioned in (2) above,
  for finding optimal DR-plans by defining a so-called \dfn{canonical}
  DR-plan and showing a strong Church-Rosser property: \vemph{all
  canonical DR-plans for isostatic or under-constrained 2D qusecs are
  optimal} (Section \ref{sec:DRP}).

@@

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system. Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems \uncited. When the required
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in \uncited\uncited (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems \uncited)
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems).

### 2014-12-18 doc/GMP2015/s03_drp.tex

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \emph{assume} is well-constrained. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be well-constrained. As $C$ is well-constrained, this means no proper subset of the children in $CompleteDRP(C)$ can union to form $C$.

@@

However, to guarantee that any two are \emph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant. Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of this node will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
the children of this node will be
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
and
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once. Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
\end{enumerate}

@@

% This is general, not for 2d
\begin{remark}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that:
\begin{enumerate}
    \item $F_i\cup F_j$ is not trivial.
    \item $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial.
    \item $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained.
    \item $F_i\cap F_j$ is not under-constrained.
\end{enumerate}
\end{remark}

\begin{proof}
Note that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
\begin{enumerate}
    \item By definition of trivial graph; if $F_i\cup F_j$ were trivial, then $F_i$ and $F_j$ must be trivial.

    \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$. Therefore $d(F_i\cap F_j)=2k-l>k$. This means $F_i\cap F_j$ is trivial. \textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial. By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.

    \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$. Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained. \textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.

    \item Subgraphs of well-constrained graph $F$ can only be trivial, under-constrained, or well-constrained. So it follows from the above that an under-constrained intersection is not possible.
\end{enumerate}
\end{proof}

@@

We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.

\begin{figure*}\centering
\begin{subfigure}{.3\linewidth}\centering
    \input{img/tikz/two_doublets-0.tikz}
    \caption{}\label{fig:doublets:a}
\end{subfigure}%
\begin{subfigure}{.7\linewidth}\centering
    \input{img/tikz/two_doublets-1.tikz}
    \caption{}\label{fig:doublets:b}
\end{subfigure}

\caption{(\ref{fig:doublets:a}) Two doublets ($C2\times C3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$. (\ref{fig:doublets:b}) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
\label{fig:doublets}
\end{figure*}

\begin{example}
    In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan. These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$. This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 2. Also shown is the decomposition of the individual doublets. These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them. As explained in Theorem \ref{theorem:main}, point 1, all of them must be included in the optimal DR-plan. Were any left out, the union of the children would not be the entire parent.
\end{example}

@@

\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K3,3$. This pattern continues for a total of $N$ triangles. (\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K3,3$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K3,3$ subgraphs. The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
\label{fig:c2c3ofk33s}
\end{figure*}

@@

\subsection{Algorithm}
The algorithm to find the optimal DR-plan is a straightforward application of the theory. It begins by finding the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs.

>>>>

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be well-constrained. As $C$ is well-constrained, this means no proper subset of the children in $CompleteDRP(C)$ can union to form $C$.

@@

However, to guarantee that any two are \vemph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant. Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of this node will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
the children of this node will be
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
and
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once. Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
\end{enumerate}

@@

% This is general, not for 2d
\begin{remark}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are well-constrained subgraphs of the same well-constrained graph $F$, then it must be that: (1) $F_i\cup F_j$ is not trivial; (2) $F_i\cup F_j$ is under-constrained if and only if $F_i\cap F_j$ is trivial; (3) $F_i\cup F_j$ is well-constrained if and only if $F_i\cap F_j$ is well-constrained; and (4) $F_i\cap F_j$ is not under-constrained.
\end{remark}

\begin{proof}
For (1), simply note that if $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.

For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$. For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.

For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
% \begin{enumerate}
%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.

%     \item Observe that  Then it is straightforward application of the equation above.

%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.

%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.

%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$. Therefore $d(F_i\cap F_j)=2k-l>k$. This means $F_i\cap F_j$ is trivial. \textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial. By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.

%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$. Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained. \textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.

%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
% \end{enumerate}
\end{proof}

@@

\todo{Explain just second figure now.}
% We now present two examples of non-tree-decomposable graphs that illustrate the the strength and flexibility of the canonical DR-plan.

% \begin{figure*}\centering
% \begin{subfigure}{.3\linewidth}\centering
%     \input{img/tikz/two_doublets-0.tikz}
%     \caption{}\label{fig:doublets:a}
% \end{subfigure}%
% \begin{subfigure}{.7\linewidth}\centering
%     \input{img/tikz/two_doublets-1.tikz}
%     \caption{}\label{fig:doublets:b}
% \end{subfigure}

% \caption{(\ref{fig:doublets:a}) Two doublets ($C_2\times C_3$), $\{a_1,a_2,a_3,b_1,b_2,b_3\}$ and $\{b_1,b_2,b_3,c_1,c_2,c_3\}$, intersecting on the triangle $\{b_1,b_2,b_3\}$. (\ref{fig:doublets:b}) The optimal DR-plan of the graph if we consider it to be a 2D graph; omits further decomposition of the three triangles into their edges and of edges into their individual nodes.}
% \label{fig:doublets}
% \end{figure*}

% \begin{example}
%     In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan. These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$. This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 1. Also shown is the decomposition of the individual doublets. These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them. As explained in Theorem \ref{theorem:main}, point 2, all of them must be included in the optimal DR-plan. Were any left out, the union of the children would not be the entire parent.
% \end{example}

@@

\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues for a total of $N$ triangles. (\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
\label{fig:c2c3ofk33s}
\end{figure*}

@@

\subsection{Algorithm}

\todo{Update, explain pebble game and frontier, complexity}

The algorithm to find the optimal DR-plan is a straightforward application of the theory. It begins by finding the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs.
